---
title: "Analisis of individual Questions EKE Colombia - Banano and Plantain 2023"
author: "Jacobo Robledo"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries, Functions, and Global Variables

#### `tree_map_3` Function

**Description:**\
This function creates a treemap visualization based on a given dataset.
It first calculates the frequency of occurrences for specified grouping variables and then plots a treemap using these frequencies.

**Parameters:**\

-   `data`: A data frame containing the data to be visualized.

-   `group`: The name of the column in 'data' to be used as the primary grouping variable.

-   `subgroup_1`: The name of the column in 'data' to be used as the first level subgroup.

-   `subgroup_2`: The name of the column in 'data' to be used as the second level subgroup.

-   `color_column`: The name of the column in 'data' that determines the color of the treemap blocks.

-   `title`: The title to be displayed on the treemap.

**Details:**\
The function groups the data by 'group', 'subgroup_1', 'subgroup_2', and 'color_column', then calculates the frequency of each group.
It uses the 'treemap' package to create the treemap, where the size of each block is determined by the calculated frequency.
The color of the blocks is determined by the 'color_column'.
The function allows customization of label sizes, colors, and placement, as well as border colors and widths.

**Returns:**\
A treemap plot as specified by the input parameters.

**Example Usage:**\
`tree_map_3(my_data, "main_category", "sub_category1", "sub_category2", "color_var", "My Treemap Title")`

#### `Col_map_by_depto` Function

**Description:**\
This function creates a colored map visualization based on departmental data.
It utilizes the `colmap` function to map data to specific geographical areas (departments) and applies a continuous color scale to these areas.

**Function Structure:** - `colmap(departamentos, data, data_id, var)`: Maps the `var` variable from the `data` data frame onto the `departamentos` geographical layout.
The `data_id` specifies the matching column in `data` that corresponds to department IDs.
- `scale_fill_continuous(low, high, na.value)`: Applies a continuous color scale to the map.
The `low` and `high` parameters define the color range for the lowest and highest values of `var`, respectively.
The `na.value` parameter defines the color for missing values.
- `labs(fill = legend_lab)`: Sets the label for the map legend.

**Parameters:**\
- `departamentos`: The geographical layout of departments.
- `data`: The data frame containing the data to be visualized.
- `data_id`: The column in `data` that corresponds to department IDs.
- `var`: The variable in `data` to be visualized on the map.
- `low`: The color representing the low end of the `var` variable's range.
- `high`: The color representing the high end of the `var` variable's range.
- `na.value`: The color used for missing values in `var`.
- `legend_lab`: The label for the map's legend.

**Example Usage:**\
Suppose you have a data frame `my_data` with department IDs in `depto_id` column and a variable `population` to visualize.
You could call:

`R Col_map_by_depto(departamentos = my_map_layout, data = my_data, data_id = "depto_id", var = "population", low = "blue", high = "red", na.value = "grey", legend_lab = "Population")`

```{r libraries}
#libraries----
  library(treemap)
  library(tidyverse)
  library(RColorBrewer)
  library("colmaps")
  library("homicidios")
#Global variables
  #set the color palette for bananas and plantains
  palette_banana <- colorRampPalette(c("#FFDA00","#FFF2CC" ))
  palette_plantain <- colorRampPalette(c("#28B463","#BEF4BE"))
  #data set: individual questions:
  #read individual surveys from GitHub
  num_cols <- length(read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/DATA_Individual_surveys_Banana_and_Plantain_Colombia_Clean_surveys_2023-12-28.csv", nrows = 1, header = TRUE))
  # Create a colClasses vector with "character" for each column
  col_classes <- rep("character", num_cols)
  individual_surveys<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/DATA_Individual_surveys_Banana_and_Plantain_Colombia_Clean_surveys_2023-12-28.csv", colClasses = col_classes)
  #reformat columns
  individual_surveys$question_number<- as.integer(individual_surveys$question_number)


#function tree_map_3 
  tree_map_3<- function(data,group, subgroup_1, subgroup_2, color_column, title){
    #Calculate the frequencies
    frequency_personalized_question<- data %>%
      group_by(!!rlang::sym(group), !!rlang::sym(subgroup_1), !!rlang::sym(subgroup_2), !!rlang::sym(color_column)) %>%
      summarise(frequency=n())
    #Plot the tree map 
    figure<- treemap(frequency_personalized_question,
            index=c(group,subgroup_1, subgroup_2), 
            vSize="frequency",
            vColor=color_column,
            type="color",
            fontsize.labels=c(0,7,4),                # size of labels. Give the size per level of aggregation: size for group, size for subgroup, sub-subgroups...
            fontcolor.labels=c("transparent","#654321", "#A58F65"),    # Color of labels
            fontface.labels=c(2,1,1),                  # Font of labels: 1,2,3,4 for normal, bold, italic, bold-italic...
            bg.labels=c("transparent"),              # Background color of labels
            align.labels=list(
              c("center", "top"), 
              c("right", "bottom"), 
              c("left", "top")
            ),                                   # Where to place labels in the rectangle?
            overlap.labels=0.5,                      # number between 0 and 1 that determines the tolerance of the overlap between labels. 0 means that labels of lower levels are not printed if higher level labels overlap, 1  means that labels are always printed. In-between values, for instance the default value .5, means that lower level labels are printed if other labels do not overlap with more than .5  times their area size.
            inflate.labels=T,                        # If true, labels are bigger when rectangle is bigger.
            border.col=c("white","#654321","#A58F65"),             # Color of borders of groups, of subgroups, of subsubgroups ....
            border.lwds=c(3,2, 0.5),                         # Width of colors
            title=title,                      # Customize your title
            #fontsize.title=12                       # Size of the title
    )
    return(figure)
  }

#function Col_map_by_depto
  Col_map_by_depto<- function(data, data_id, var, low, high, na.values, legend_lab){
    
    figure<- colmap(departamentos, data = data, data_id = data_id, var = var)+
      scale_fill_continuous(low = low , high = high, na.value = na.values)
    labs(fill= legend_lab)
    return(figure)
  }
```

## Question 1: `r individual_surveys$question_in_english[1]`

```{r question_1}
#selecting only question 1----
  personalized_question<- filter(individual_surveys, question_number==1)
  #organizing the order of the levels
  personalized_question$question_in_english<- factor(personalized_question$question_in_english,levels = c("1 to 2", "3 to 5","5 to 9","10 to 15","More than 15"))
  #finding the number of levels per crop (banana and plantain)
  levels_per_crop_Q1<- tapply(personalized_question$answer_in_english, personalized_question$crop, function(x){length(unique(x))})
  #generating a ramp palette according to the number of levels per crop
  colors_banana_question1<- palette_banana(levels_per_crop_Q1["Banana"])
  colors_plantain_question1<- palette_plantain(levels_per_crop_Q1["Plantain"])
  #Assigning the colors by crop and answer
  personalized_question<- personalized_question %>%
    mutate(group_color= case_when(
      crop=="Banana" & answer_in_english == "1 to 2" ~ colors_banana_question1[5],
      crop=="Banana" & answer_in_english == "3 to 5" ~ colors_banana_question1[4],
      crop=="Banana" & answer_in_english == "5 to 9" ~ colors_banana_question1[3],
      crop=="Banana" & answer_in_english == "10 to 15" ~ colors_banana_question1[2],
      crop=="Banana" & answer_in_english == "More than 15" ~ colors_banana_question1[1],
      crop=="Plantain" & answer_in_english == "1 to 2" ~ colors_plantain_question1[5],
      crop=="Plantain" & answer_in_english == "3 to 5" ~ colors_plantain_question1[4],
      crop=="Plantain" & answer_in_english == "5 to 9" ~ colors_plantain_question1[3],
      crop=="Plantain" & answer_in_english == "10 to 15" ~ colors_plantain_question1[2],
      crop=="Plantain" & answer_in_english == "More than 15" ~ colors_plantain_question1[1],
    ))
#tree map
tree_map_3(data = personalized_question, "crop", "expert_in", "answer_in_english", "group_color", title="Experience of Experts by Department")
#summarise to have the number of experts by departments
data<- personalized_question %>%
  group_by(expert_in, id_depto) %>%
  summarise(Number_of_Experts= n())
codes_department<- unique(data.frame(EKE.expert.in=homicidios$depto, id_depto=homicidios$id_depto))
data<- full_join(codes_department, data)


Col_map_by_depto(data = data, data_id = "id_depto", var = "Number_of_Experts", low = "#A58F65", high = "#654321", na.values = "#eeeeee", legend_lab = "Number \n of experts")
```
