---
title: "Analisis of individual Questions EKE Colombia - Banano and Plantain 2023"
author: "Jacobo Robledo"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries, Functions, and Global Variables

#### `tree_map_3` Function

**Description:**\
This function creates a treemap visualization based on a given dataset.
It first calculates the frequency of occurrences for specified grouping variables and then plots a treemap using these frequencies.

**Parameters:**\

-   `data`: A data frame containing the data to be visualized.

-   `group`: The name of the column in 'data' to be used as the primary grouping variable.

-   `subgroup_1`: The name of the column in 'data' to be used as the first level subgroup.

-   `subgroup_2`: The name of the column in 'data' to be used as the second level subgroup.

-   `color_column`: The name of the column in 'data' that determines the color of the treemap blocks.

-   `title`: The title to be displayed on the treemap.

**Details:**\
The function groups the data by 'group', 'subgroup_1', 'subgroup_2', and 'color_column', then calculates the frequency of each group.
It uses the 'treemap' package to create the treemap, where the size of each block is determined by the calculated frequency.
The color of the blocks is determined by the 'color_column'.
The function allows customization of label sizes, colors, and placement, as well as border colors and widths.

**Returns:**\
A treemap plot as specified by the input parameters.

**Example Usage:**\
`tree_map_3(my_data, "main_category", "sub_category1", "sub_category2", "color_var", "My Treemap Title")`

#### `Col_map_by_depto` Function

**Description:**\
This function creates a colored map visualization based on departmental data.
It utilizes the `colmap` function to map data to specific geographical areas (departments) and applies a continuous color scale to these areas.

**Function Structure:** - `colmap(departamentos, data, data_id, var)`: Maps the `var` variable from the `data` data frame onto the `departamentos` geographical layout.
The `data_id` specifies the matching column in `data` that corresponds to department IDs.
- `scale_fill_continuous(low, high, na.value)`: Applies a continuous color scale to the map.
The `low` and `high` parameters define the color range for the lowest and highest values of `var`, respectively.
The `na.value` parameter defines the color for missing values.
- `labs(fill = legend_lab)`: Sets the label for the map legend.

**Parameters:**\
- `departamentos`: The geographical layout of departments.
- `data`: The data frame containing the data to be visualized.
- `data_id`: The column in `data` that corresponds to department IDs.
- `var`: The variable in `data` to be visualized on the map.
- `low`: The color representing the low end of the `var` variable's range.
- `high`: The color representing the high end of the `var` variable's range.
- `na.value`: The color used for missing values in `var`.
- `legend_lab`: The label for the map's legend.

**Example Usage:**\
Suppose you have a data frame `my_data` with department IDs in `depto_id` column and a variable `population` to visualize.
You could call:

`R Col_map_by_depto(departamentos = my_map_layout, data = my_data, data_id = "depto_id", var = "population", low = "blue", high = "red", na.value = "grey", legend_lab = "Population")`

#### `category_by_mean_by_question` Function

**Description:**\
This function retrieves the category corresponding to a given mean value for a specific question.
It reads a dataset from a specified URL containing category information in both English and Spanish.
The function allows the user to specify the language for the category description.

**Parameters:**\
- `question`: The question number for which the category needs to be found.
- `mean`: The mean value whose corresponding category is to be retrieved.
- `language`: A character vector specifying the language of the category description.
Default is `c("en", "es")`.

**Functionality:**\
- The function first rounds the mean value to the nearest whole number.
- It then reads a dataset from a given URL that contains categories in both English and Spanish.
- Depending on the selected language, it filters the dataset to find the category corresponding to the provided question number and rounded mean value.
- If the language is set to "en" (English), it retrieves the category from the `trans_answer` column; if "es" (Spanish), it retrieves from the `Category` column.
- An error message is displayed if the specified language is not found.

**Returns:**\
The category corresponding to the specified mean and question in the chosen language.

**Example Usage:**\
Suppose you have a mean value of 3.6 for question number 5 and want the category in English.
You could call:

`R   category <- category_by_mean_by_question(question = 5, mean = 3.6, language = "en")`

#### `count_elements_by_group` Function

**Description:**\
This function is designed to count the frequency of elements within a specified column that contains comma-separated values.
It processes the data to convert it into a long format and then performs a count of each element within the specified groups.

**Parameters:**\
- `data`: The data frame containing the data to be analyzed.
- `value_column`: The name of the column in `data` that contains the comma-separated values.
- `group_columns`: A vector of column names in `data` used to define groups for aggregation.

**Functionality:**\
- The function first uses `separate_rows` to split the comma-separated values in `value_column` and convert the data into a long format.
- `group_by` is then applied to group the data by the columns specified in `group_columns`.
- `count` is used to count the occurrences of each unique element within these groups.
- The result is a long-format data frame with the frequency count of each element in the specified groups.

**Returns:**\
A data frame in long format that contains the counts of each unique element within the specified groups.

**Example Usage:**\
Suppose you have a data frame `df` with a column `categories` containing comma-separated values and you want to count these values within groups defined by `column1` and `column2`.
You could call: `R   result <- count_elements_by_group(df, "categories", c("column1", "column2"))`

#### `bar_plot_banana_plantain` Function

**Description:**\
This function creates a bar plot specifically designed for visualizing data related to 'Banana' and 'Plantain' categories.
It includes customizable aesthetics such as background color for highlighting, bar color coding, and an adjustable upper limit for the y-axis to enhance the proportional representation of data.

**Parameters:**\
- `data_set`: The data frame containing the data to be plotted.
- `x_lab`: Label for the x-axis.
- `y_lab`: Label for the y-axis.
- `title`: The title of the plot.
- `x`: The name of the column in `data_set` to be used as the x-axis variable.
- `y`: The name of the column in `data_set` to be used as the y-axis variable.
- `facet`: The name of the column in `data_set` to create facets for different panels in the plot.
- `alpha`: The transparency level for the background color of the geom_rect.
- `background_color`: The name of the column in `data_set` that contains the background colors for the rectangles.
- `bar_color`: The name of the column in `data_set` that contains the colors for the bars.
- `proportional_limit`: A numeric value to set the upper limit of the y-axis as a proportion of the maximum value of the y variable.

**Functionality:**\
- The function starts by converting provided column names into symbols for `ggplot2` aesthetics.
- It calculates the new y-axis upper limit based on the provided `proportional_limit`.
- A ggplot object is initialized, and layers are added for rectangles, text, and bars, with aesthetics mapped accordingly.
- The plot's appearance is customized, including colors for text and lines, and removing the grid lines for a cleaner look.
- The y-axis is scaled with the calculated limit, and facet panels are created as per the `facet` column.

**Returns:**\
A ggplot object representing the bar plot with customized aesthetics and scales, ready for rendering or further modification.

**Example Usage:**\
To create a plot for a data frame `df` with background colors specified in `bg_colors` and bar colors in `bar_colors`, where we want to facet by the 'type' column and set the y-axis upper limit to be 20% higher than the max 'value' in `df`:

`R result_plot <- bar_plot_banana_plantain(   data_set = df,   x_lab = "Type of Crop",   y_lab = "Value in Percentage",   title = "Crop Value Comparison",   x = "type",   y = "value",   facet = "category",   alpha = 0.3,   background_color = "bg_colors",   bar_color = "bar_colors",   proportional_limit = 1.2 ) # Print the result print(result_plot)`

#### `assign_group_color` Function

**Description:**\
This function dynamically assigns color codes to data points based on their categorical responses within specific groups (crops).
It is designed to work with variable-length categories and color mappings for different groups.
It's especially useful for visualizing data where the color representation depends on both the group (crop) and the category of response.

**Parameters:**\
- `data`: The data frame containing the data to be processed.
- `levels_by_crop`: A named list where each element is a vector of categorical levels for a specific crop.
- `colors_by_crop`: A named list where each element is a vector of colors corresponding to each categorical level for each crop.

**Functionality:**\
- The function iterates through each row of the data frame.
- For each row, it matches the crop and its corresponding answer to its categorical level.
- It then assigns the color based on the matched level from the provided color vectors.
- The function handles varying lengths of categories and color mappings, making it versatile for different datasets.

**Returns:**\
The modified data frame with an additional column `group_color` that contains the assigned color codes based on the crop and answer categories.

**Example Usage:**\
To assign colors to a dataset `df` with crops 'Banana' and 'Plantain' and their respective levels and colors:

`r levels_by_crop <- list( Banana = c("Very Low Knowledge", "Low Knowledge", "Moderate Knowledge", "High Knowledge", "Very High Knowledge"), Plantain = c("No Knowledge", "Very Low Knowledge", "Low Knowledge", "Moderate Knowledge", "High Knowledge", "Very High Knowledge") )`


`r colors_by_crop <- list( Banana = c("#color1", "#color2", "#color3", "#color4", "#color5"), Plantain = c("#color6", "#color7", "#color8", "#color9", "#color10", "#color11") )`


`r #df <- assign_group_color(df, levels_by_crop, colors_by_crop)`



```{r libraries}
#libraries----
  library(treemap)
  library(tidyverse)
  library(RColorBrewer)
  library("colmaps")
  library("homicidios")
  library(voronoiTreemap)
  library(webshot)
  library(webshot2)
  library(purrr)
  library(rlang)

#Global variables
  #set the color palette for bananas and plantains
  palette_banana <- colorRampPalette(c("#FFDA00","#FFF2CC" ))
  palette_plantain <- colorRampPalette(c("#28B463","#BEF4BE"))
  palette_soil<- colorRampPalette(c( "#654321","#d8c298"))
  #data set: individual questions:
  #read individual surveys from GitHub
  num_cols <- length(read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/DATA_Individual_surveys_Banana_and_Plantain_Colombia_Clean_surveys_2023-12-28.csv", nrows = 1, header = TRUE))
  # Create a colClasses vector with "character" for each column
  col_classes <- rep("character", num_cols)
  individual_surveys<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/DATA_Individual_surveys_Banana_and_Plantain_Colombia_Clean_surveys_2023-12-28.csv", colClasses = col_classes)
  #reformat columns
  individual_surveys$question_number<- as.integer(individual_surveys$question_number)
  #Getting the experience of each expert by department by crop
  experience_of_each_expert<- individual_surveys %>%
    filter(question_number==1) %>%
    select(crop, expert_ID, expert_in, numeric_answer) %>%
    mutate(numeric_answer= as.numeric(numeric_answer))
  colnames(experience_of_each_expert)[which(colnames(experience_of_each_expert)=="numeric_answer")]<- "weights"
  #correct the missing answer which is "Antioquia"
  experience_of_each_expert$expert_in[which(is.na(experience_of_each_expert$expert_in))]<- "Antioquia"
    
#function tree_map_3 
  tree_map_3<- function(data,group, subgroup_1, subgroup_2, color_column, title){
    #Calculate the frequencies
    frequency_personalized_question<- data %>%
      group_by(!!rlang::sym(group), !!rlang::sym(subgroup_1), !!rlang::sym(subgroup_2), !!rlang::sym(color_column)) %>%
      summarise(frequency=n())
    #Plot the tree map 
    figure<- treemap(frequency_personalized_question,
            index=c(group,subgroup_1, subgroup_2), 
            vSize="frequency",
            vColor=color_column,
            type="color",
            fontsize.labels=c(0,7,4),                # size of labels. Give the size per level of aggregation: size for group, size for subgroup, sub-subgroups...
            fontcolor.labels=c("transparent","#654321", "#A58F65"),    # Color of labels
            fontface.labels=c(2,1,1),                  # Font of labels: 1,2,3,4 for normal, bold, italic, bold-italic...
            bg.labels=c("transparent"),              # Background color of labels
            align.labels=list(
              c("center", "top"), 
              c("right", "bottom"), 
              c("left", "top")
            ),                                   # Where to place labels in the rectangle?
            overlap.labels=0.5,                      # number between 0 and 1 that determines the tolerance of the overlap between labels. 0 means that labels of lower levels are not printed if higher level labels overlap, 1  means that labels are always printed. In-between values, for instance the default value .5, means that lower level labels are printed if other labels do not overlap with more than .5  times their area size.
            inflate.labels=T,                        # If true, labels are bigger when rectangle is bigger.
            border.col=c("white","#654321","#A58F65"),             # Color of borders of groups, of subgroups, of subsubgroups ....
            border.lwds=c(3,2, 0.5),                         # Width of colors
            title=title,                      # Customize your title
            #fontsize.title=12                       # Size of the title
    )
    return(figure)
  }

#function Col_map_by_depto
  Col_map_by_depto<- function(data, data_id, var, low, high, na.values, legend_lab){
    
    figure<- colmap(departamentos, data = data, data_id = data_id, var = var)+
      scale_fill_continuous(low = low , high = high, na.value = na.values)
    labs(fill= legend_lab)
    return(figure)
  }
  
#function category_by_mean_by_question
category_by_mean_by_question<- function(question, mean, language=c("en","es")){
  mean<-round(mean, 0)
  data<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/SUP_DATA_Categories_that_can_be_chosen_in_every_question_of_the_individual_survey_English_and_Spanish.csv")
  if (language=="en") {
    category<- filter(data, number==question & number_to_category==mean)$trans_answer
  }else{
    if (language=="es") {
      category<- filter(data, number==question & number_to_category==mean)$Category
    }
    cat("ERROR: Language not found")
  }
  return(category)
}

#Function count_elements_by_group
count_elements_by_group <- function(data, value_column, group_columns) {
  # Split the comma-separated values and convert to long format
  data_long <- data %>%
    separate_rows(!!sym(value_column), sep = ",\\s*") %>%
    group_by(across(all_of(group_columns))) %>%
    count(!!sym(value_column))
  return(data_long)
}
  
#function bar_plot_banana_plantain
bar_plot_banana_plantain <- function(data_set, x_lab, y_lab, title, x, y, alpha, background_color, bar_color, proportional_limit, facet = NULL) {
  # Calculate new y-axis upper limit, 20% above the max value
  ymax_limit <- max(data_set[[y]], na.rm = TRUE) * proportional_limit
  
  # Convert strings to symbols
  x_sym <- rlang::sym(x)
  y_sym <- rlang::sym(y)
  background_color_sym <- rlang::sym(background_color)
  bar_color_sym <- rlang::sym(bar_color)
  
  # Start the plot
  p <- ggplot(data_set, aes(x = !!x_sym, y = !!y_sym)) +
    geom_rect(aes(fill = !!background_color_sym), xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, alpha = alpha) +
    geom_text(aes(label = !!y_sym), vjust = -0.5, color = "#654321") +
    geom_bar(aes(fill = !!bar_color_sym), color = "#654321",stat = 'identity', position = position_dodge()) +
    scale_fill_identity() +
    labs(x = x_lab, y = y_lab, title = title) +
    theme(
      text = element_text(color = "#654321"),
      axis.title = element_text(color = "#654321"),
      axis.text = element_text(color = "#654321"),
      axis.line = element_line(color = "#654321"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "#654321", fill = NA),
      strip.text = element_text(color = "#654321"),
      strip.background = element_rect(fill = "#A58F65", colour = NA),
      legend.position = "none"
    ) +
    ylim(0, ymax_limit)
  
  # Conditionally add facet grid if facet is provided
  if (!is.null(facet) && facet != "") {
    facet_sym <- rlang::sym(facet)
    facet_formula <- as.formula(paste(". ~", facet))
    p <- p + facet_grid(facet_formula)
  }
  
  # Return the plot
  return(p)
}

#function assign_group_color 
assign_group_color <- function(data, answer_col, levels_by_crop, colors_by_crop) {
  # Ensure that answer_col is a string representing the column name
  # levels_by_crop is a named list of vectors containing the knowledge levels for each crop
  # colors_by_crop is a named list of vectors containing the colors corresponding to each knowledge level for each crop
  
  answer_col_sym <- rlang::sym(answer_col)  # Convert the string to a symbol
  
  data <- data %>%
    mutate(group_color = purrr::map2_chr(crop, !!answer_col_sym, function(crop_name, answer) {
      # Find the index of the answer within the levels for the given crop
      level_index <- match(answer, levels_by_crop[[crop_name]])
      # Use the index to get the corresponding color
      if (!is.na(level_index)) {
        colors_by_crop[[crop_name]][level_index]
      } else {
        NA_character_  # If there's no match, return NA
      }
    }))
  
  return(data)
}
```

## Question 1: `r unique(individual_surveys$question_in_english)[1]`

```{r question_1}
#selecting only question 1----
  question_1<- filter(individual_surveys, question_number==1)
  #organizing the order of the levels
  question_1$answer_in_english<- factor(question_1$answer_in_english,levels = c("1 to 2", "3 to 5","5 to 9","10 to 15","More than 15"))
  #finding the number of levels per crop (banana and plantain)
  levels_per_crop_Q1<- tapply(question_1$answer_in_english, question_1$crop, function(x){length(unique(x))})
  #generating a ramp palette according to the number of levels per crop
  colors_banana_question1<- palette_banana(levels_per_crop_Q1["Banana"])
  colors_plantain_question1<- palette_plantain(levels_per_crop_Q1["Plantain"])
  #Assigning the colors by crop and answer
  question_1<- question_1 %>%
    mutate(group_color= case_when(
      crop=="Banana" & answer_in_english == "1 to 2" ~ colors_banana_question1[5],
      crop=="Banana" & answer_in_english == "3 to 5" ~ colors_banana_question1[4],
      crop=="Banana" & answer_in_english == "5 to 9" ~ colors_banana_question1[3],
      crop=="Banana" & answer_in_english == "10 to 15" ~ colors_banana_question1[2],
      crop=="Banana" & answer_in_english == "More than 15" ~ colors_banana_question1[1],
      crop=="Plantain" & answer_in_english == "1 to 2" ~ colors_plantain_question1[5],
      crop=="Plantain" & answer_in_english == "3 to 5" ~ colors_plantain_question1[4],
      crop=="Plantain" & answer_in_english == "5 to 9" ~ colors_plantain_question1[3],
      crop=="Plantain" & answer_in_english == "10 to 15" ~ colors_plantain_question1[2],
      crop=="Plantain" & answer_in_english == "More than 15" ~ colors_plantain_question1[1],
    ))
#tree map
tree_map_3(data = question_1, "crop", "expert_in", "answer_in_english", "group_color", title="Experience of Experts by Department")
#summarise to have the number of experts by departments
data<- question_1 %>%
  group_by(expert_in, id_depto) %>%
  summarise(Number_of_Experts= n())
codes_department<- unique(data.frame(EKE.expert.in=homicidios$depto, id_depto=homicidios$id_depto))
data<- full_join(codes_department, data)
#Map of Colombia
Col_map_by_depto(data = data, data_id = "id_depto", var = "Number_of_Experts", low = "#A58F65", high = "#654321", na.values = "#eeeeee", legend_lab = "Number \n of experts")
#Total number of experts:
total_experts_Q1<- length(unique(question_1$expert_ID))
mode_Q1<- sort(table(question_1$answer_in_english), decreasing = T)
question_1$numeric_answer<- as.numeric(question_1$numeric_answer)
mean_q1<- mean(question_1$numeric_answer)
mean_q1<- category_by_mean_by_question(1, mean_q1, language = "en")
mean_q1_by_crops<- question_1 %>%
  group_by(crop) %>%
  summarise(mean_cat= mean(numeric_answer),
            surveys_by_crop= length(unique(expert_ID)),
            mode= names(sort(table(answer_in_english), decreasing = T))[1])
mean_q1_by_crops$cat_english<- tapply(mean_q1_by_crops$mean_cat, mean_q1_by_crops$crop, function(x){
  category_by_mean_by_question(1, mean = x, language = "en")}) 
  
```

#### Descriptive statistics:

Total number of experts that answer this question= `r total_experts_Q1`

Mode in all the surveys= `r names(mode_Q1[1])`

Mean in all the surveys=`r mean_q1`

Descriptive statistics by crop=

```{r stats_q1, echo=FALSE }
library(knitr)
kable(mean_q1_by_crops)
```

## Question 2: `r unique(individual_surveys$question_in_english)[2]`

```{r question_2}
question_2<- filter(individual_surveys,question_number==2)
question_2_frequencies <- count_elements_by_group(question_2, "answer_in_english", c("crop"))
question_2_frequencies$answer_in_english<- as.factor(question_2_frequencies$answer_in_english)
#finding the number of levels per crop (banana and plantain)
levels_per_crop_Q2<- tapply(question_2_frequencies$answer_in_english, question_2_frequencies$crop, function(x){length(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question2<- palette_banana(levels_per_crop_Q2["Banana"])
colors_plantain_question2<- palette_plantain(levels_per_crop_Q2["Plantain"])
#Assignig a color per factor by crop
question_2_frequencies <- question_2_frequencies %>%
  mutate(group_color = case_when(
    crop == "Banana" ~ colors_banana_question2[match(answer_in_english, c("Agricultural Extension", 
                                                                          "Agricultural Operations Management/Crop Consultants", 
                                                                          "Agronomy", "Economics", 
                                                                          "Entomology", "Horticulture", 
                                                                          "IPM (Integrated Pest Management)", 
                                                                          "Others", "Plant Pathology", 
                                                                          "Producer", "Research", 
                                                                          "Seed Systems", "Social Sciences"))],
    crop == "Plantain" ~ colors_plantain_question2[match(answer_in_english, c("Agricultural Extension", 
                                                                              "Agricultural Operations Management/Crop Consultants", 
                                                                              "Agronomy", "Economics", 
                                                                              "Entomology", "Horticulture", 
                                                                              "IPM (Integrated Pest Management)", 
                                                                              "Others", "Plant Pathology", 
                                                                              "Producer", "Research", 
                                                                              "Seed Systems", "Social Sciences"))]
  ))
#Creating avreviations for the the plot 
question_2_frequencies <- question_2_frequencies %>%
  mutate(cat_abbreviations = case_when(
    answer_in_english == "Agricultural Extension" ~ "AE",
    answer_in_english == "Agricultural Operations Management/Crop Consultants" ~ "AOM/CC",
    answer_in_english == "Agronomy" ~ "Agro",
    answer_in_english == "Economics" ~ "Econ",
    answer_in_english == "Entomology" ~ "Ento",
    answer_in_english == "Horticulture" ~ "Hort",
    answer_in_english == "IPM (Integrated Pest Management)" ~ "IPM",
    answer_in_english == "Others" ~ "Oth",
    answer_in_english == "Plant Pathology" ~ "PlPath",
    answer_in_english == "Producer" ~ "Prod",
    answer_in_english == "Research" ~ "Res",
    answer_in_english == "Seed Systems" ~ "SS",
    answer_in_english == "Social Sciences" ~ "SocSci",
    TRUE ~ NA_character_  # Default case if none of the above conditions are met
  ))
#formating the data set to make the visualization, voronoiTreemap was quite picky...  
colnames(question_2_frequencies)<- c('h2', "h3", "weight","color", "codes")
question_2_frequencies$h1<- "Total"
question_2_frequencies <- question_2_frequencies %>%
  select(h1, h2, h3, color, weight, codes)
question_2_frequencies$weight<- (question_2_frequencies$weight/sum(question_2_frequencies$weight))*100
question_2_frequencies$h1<- as.factor(question_2_frequencies$h1)
question_2_frequencies$h2<- as.factor(question_2_frequencies$h2)
question_2_frequencies<- data.frame(h1= question_2_frequencies$h1,
                                    h2= question_2_frequencies$h2,
                                    h3= question_2_frequencies$h3,
                                    color= question_2_frequencies$color,
                                    weight= question_2_frequencies$weight,
                                    codes= question_2_frequencies$codes)
gdp_json <- vt_export_json(vt_input_from_df(question_2_frequencies))
vt_d3(gdp_json, color_border = "#654321", size_border = "1.5px", label = T, color_label = "#654321", seed = 3)
vt_d3(gdp_json, color_border = "#654321", size_border = "1.5px", label = F, color_label = "#654321", seed = 3)
kable(question_2_frequencies)
```

## Question 3: `r unique(individual_surveys$question_in_english)[3]`

```{r queestion_3}
#selecting only question 3----
  question_3<- filter(individual_surveys, question_number==3)
#Counts by crop
  question_3_frequencies <- count_elements_by_group(question_3, "answer_in_english", c("crop"))
  #calculate the percentages by crop and question 
  question_3_frequencies <- question_3_frequencies %>%
  group_by(crop) %>%
  mutate(total = sum(n),          # Calculate total for each crop group
         percentage = (n / total) * 100) %>%
  ungroup()
  #Plot
  #Add colors to each 
  question_3_frequencies<- question_3_frequencies %>%
  mutate(background_color= case_when(crop=="Banana" ~ "#FFF2CC",
                     crop=="Plantain" ~ "#BEF4BE"))
  #bars colors
question_3_frequencies<- question_3_frequencies %>%
  mutate(bar_color= case_when(crop=="Banana" & answer_in_english=="Public" ~ "#FFDA00",
                              crop=="Banana" & answer_in_english=="Private" ~ "#FFEDA3",
                              crop=="Plantain" & answer_in_english=="Public" ~ "#28B463",
                              crop=="Plantain" & answer_in_english=="Private" ~ "#82DA99"))
question_3_frequencies$percentage<- round(question_3_frequencies$percentage,0)
#custom function
question_3_plot<- bar_plot_banana_plantain(
  data_set = question_3_frequencies,
  x_lab = "Sector Affiliation of Experts",
  y_lab = "Percent (%)",
  title = "Expert Affiliation by Crop",
  x = "answer_in_english",
  y = "percentage",
  facet = "crop",
  alpha = 0.3,
  background_color = "background_color",
  bar_color = "bar_color",
  proportional_limit= 1.05
)
question_3_plot
#Counts national
  question_3_frequencies_nal <- count_elements_by_group(question_3, "answer_in_english", c("answer_in_english"))
  #calculate the percentages by crop and question 
  question_3_frequencies_nal <- question_3_frequencies_nal %>%
     ungroup() %>%
    mutate(total = sum(n),          # Calculate total for each crop group
         percentage = (n / total) * 100) 
  #Plot
  #Add colors to each 
  question_3_frequencies_nal<- question_3_frequencies_nal %>%
  mutate(background_color= case_when(answer_in_english=="Public" ~ "transparent",
                     answer_in_english=="Private" ~ "transparent"))
  #Add color to the bars
 question_3_frequencies_nal<-  question_3_frequencies_nal %>%
  mutate(bar_color= case_when(answer_in_english=="Public"~ "#654321",
                              answer_in_english=="Private"~ "#d8c298"))
question_3_frequencies_nal$percentage<- round(question_3_frequencies_nal$percentage,0)
  #custom plot 
question_3_plot_nal<- bar_plot_banana_plantain(
  data_set = question_3_frequencies_nal,
  x_lab = "Sector Affiliation of Experts",
  y_lab = "Percent (%)",
  title = "Expert Affiliation",
  x = "answer_in_english",
  y = "percentage",
  alpha = 1,
  background_color = "background_color",
  bar_color = "bar_color",
  proportional_limit= 1.05
)
question_3_plot_nal
```

## Question 4: `r unique(individual_surveys$question_in_english)[4]`

```{r question_4}
#selecting only question 4----
  question_4<- filter(individual_surveys, question_number==4)
#Counts by crop
  question_4_frequencies <- count_elements_by_group(question_4, "answer_in_english", c("crop"))
  #calculate the percentages by crop and question 
  question_4_frequencies <- question_4_frequencies %>%
  group_by(crop) %>%
  mutate(total = sum(n),          # Calculate total for each crop group
         percentage = (n / total) * 100) %>%
  ungroup()
  question_4_frequencies$answer_in_english <- factor(
  question_4_frequencies$answer_in_english, 
  levels = c("No Formal Education", "Secondary Education", "Bachelor's Degree", "Master's Degree", "Doctorate"),
  labels = c("No Ed.", "Sec.", "Bac.", "Ms.C", "Ph.D")
)
  #Plot
  #Add colors to each backgorund 
  question_4_frequencies<- question_4_frequencies %>%
  mutate(background_color= case_when(crop=="Banana" ~ "#FFF2CC",
                     crop=="Plantain" ~ "#BEF4BE"))
  #Colors of the bars
  # Identify the levels by crop
  levels_Q_4_banana<- question_4_frequencies$answer_in_english[which( question_4_frequencies$crop=="Banana")]
  levels_Q_4_plantain<- question_4_frequencies$answer_in_english[which( question_4_frequencies$crop=="Plantain")]
  #set the color. The +1 is to make a different color that background wich will be the last element in the vector
  q_4_colors_banana<- palette_banana(length(levels_Q_4_banana)+1)
  q_4_colors_plantain<- palette_plantain(length(levels_Q_4_plantain)+1)
  names(q_4_colors_banana)<- levels_Q_4_banana
  names(q_4_colors_plantain)<- levels_Q_4_plantain
    # Modify the data_set to include the bar_color
question_4_frequencies <- question_4_frequencies %>%
  mutate(bar_color = case_when(
    crop == "Banana" ~ q_4_colors_banana[answer_in_english],
    crop == "Plantain" ~ q_4_colors_plantain[answer_in_english]
  ))
question_4_frequencies$percentage<- round(question_4_frequencies$percentage, 0)
dput(unique(question_4_frequencies$answer_in_english))
#custom function
question_4_plot<- bar_plot_banana_plantain(
  data_set = question_4_frequencies,
  x_lab = "Education Level of the Experts",
  y_lab = "Percent (%)",
  title = "Eucation Level of the Experts by Crop",
  x = "answer_in_english",
  y = "percentage",
  facet = "crop",
  alpha = 0.3,
  background_color = "background_color",
  bar_color = "bar_color",
  proportional_limit= 1.05
)
question_4_plot
#Counts national
  question_4_frequencies_nal <- count_elements_by_group(question_4, "answer_in_english", c("answer_in_english"))
  #calculate the percentages by crop and question 
  question_4_frequencies_nal <- question_4_frequencies_nal %>%
     ungroup() %>%
    mutate(total = sum(n),          # Calculate total for each crop group
         percentage = (n / total) * 100) 
  #Plot
  #Add colors to each background
   question_4_frequencies_nal$answer_in_english <- factor(
  question_4_frequencies_nal$answer_in_english, 
  levels = c("No Formal Education", "Secondary Education", "Bachelor's Degree", "Master's Degree", "Doctorate"),
  labels = c("No Ed.", "Sec.", "Bac.", "Ms.C", "Ph.D")
)
  question_4_frequencies_nal$background_color<-"transparent"
   q_4_colors_soil_nal<- palette_soil(length(levels(question_4_frequencies_nal$answer_in_english)))
    names(q_4_colors_soil_nal)<- levels(question_4_frequencies_nal)
 # Modify the data_set to include the bar_color
question_4_frequencies_nal <- question_4_frequencies_nal %>%
  mutate(bar_color = q_4_colors_soil_nal[answer_in_english])
question_4_frequencies_nal$percentage<- round(question_4_frequencies_nal$percentage, 0)
  #custom plot 
question_4_plot_nal<- bar_plot_banana_plantain(
  data_set = question_4_frequencies_nal,
  x_lab = "Education Level of the Experts",
  y_lab = "Percent (%)",
  title = "Education Level of the Experts",
  x = "answer_in_english",
  y = "percentage",
  alpha = 1,
  background_color = "background_color",
  bar_color = "bar_color",
  proportional_limit= 1.05
)
question_4_plot_nal
```

## Question 5: `r unique(individual_surveys$question_in_english)[5]`

```{r question_5}
question_5<- filter(individual_surveys,question_number==5)
question_5_frequencies <- count_elements_by_group(question_5, "answer_in_english", c("crop"))
question_5_frequencies$answer_in_english<- as.factor(question_5_frequencies$answer_in_english)
#finding the number of levels per crop (banana and plantain)
levels_per_crop_Q5<- tapply(question_5_frequencies$answer_in_english, question_5_frequencies$crop, function(x){length(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question5<- palette_banana(levels_per_crop_Q5["Banana"])
colors_plantain_question5<- palette_plantain(levels_per_crop_Q5["Plantain"])
#Assignig a color per factor by crop
question_5_frequencies <- question_5_frequencies %>%
  mutate(group_color = case_when(
    crop == "Banana" ~ colors_banana_question5[match(answer_in_english, c("Banana Mosaic Disease", "Black Sigatoka", 
"Black Weevil", "Elephantiasis", "Fusarium", "Moko or Madurabiche", 
"Nematodes", "Red Spider Mite", "Scale Insect", "Scarab Beetle", 
"Whiteflies", "Yellow Sigatoka"))],
    crop == "Plantain" ~ colors_plantain_question5[match(answer_in_english, c("Banana Mosaic Disease", "Black Sigatoka", 
"Black Weevil", "Elephantiasis", "Fusarium", "Moko or Madurabiche", 
"Nematodes", "Red Spider Mite", "Scale Insect", "Scarab Beetle", 
"Whiteflies", "Yellow Sigatoka"))]
  ))
#Make the abbreviations
question_5_frequencies <- question_5_frequencies %>%
  mutate(cat_abbreviations = case_when(
    answer_in_english == "Banana Mosaic Disease" ~ "BMD",
    answer_in_english == "Black Sigatoka" ~ "BS",
    answer_in_english == "Black Weevil" ~ "BW",
    answer_in_english == "Elephantiasis" ~ "ELE",
    answer_in_english == "Fusarium" ~ "FUS",
    answer_in_english == "Moko or Madurabiche" ~ "MOKO",
    answer_in_english == "Nematodes" ~ "NEM",
    answer_in_english == "Red Spider Mite" ~ "RSM",
    answer_in_english == "Scale Insect" ~ "SI",
    answer_in_english == "Scarab Beetle" ~ "SB",
    answer_in_english == "Whiteflies" ~ "WF",
    answer_in_english == "Yellow Sigatoka" ~ "YS",
    TRUE ~ NA_character_  # Default case if none of the above conditions are met
  ))

#formating the data set to make the visualization, voronoiTreemap was quite picky...  
colnames(question_5_frequencies)<- c('h2', "h3", "weight","color", "codes")
question_5_frequencies$h1<- "Total"
question_5_frequencies <- question_5_frequencies %>%
  select(h1, h2, h3, color, weight, codes)
question_5_frequencies$weight<- (question_5_frequencies$weight/sum(question_5_frequencies$weight))*100
question_5_frequencies$h1<- as.factor(question_5_frequencies$h1)
question_5_frequencies$h2<- as.factor(question_5_frequencies$h2)
question_5_frequencies<- data.frame(h1= question_5_frequencies$h1,
                                    h2= question_5_frequencies$h2,
                                    h3= question_5_frequencies$h3,
                                    color= question_5_frequencies$color,
                                    weight= question_5_frequencies$weight,
                                    codes= question_5_frequencies$codes)
gdp_json <- vt_export_json(vt_input_from_df(question_5_frequencies))
vt_d3(gdp_json, color_border = "#654321", size_border = "1.5px", label = T, color_label = "#654321", seed = 3)
vt_d3(gdp_json, color_border = "#654321", size_border = "1.5px", label = F, color_label = "#654321", seed = 3)
kable(question_5_frequencies)
#select top 5
question_5_frequencies <- question_5_frequencies %>%
  group_by(h2) %>%
  slice_max(order_by = weight, n = 5) 
kable(question_5_frequencies, label = "TOP 5 pest and Diseases by Crop (banana and plantain")
#Frequencies by department
question_5_frequencies_by_depto <- count_elements_by_group(question_5, "answer_in_english", c("crop", "expert_in"))
#colors by disease
question_5_frequencies_by_depto <- question_5_frequencies_by_depto %>%
  mutate(group_color = case_when(
    crop == "Banana" ~ colors_banana_question5[match(answer_in_english, c("Banana Mosaic Disease", "Black Sigatoka", 
"Black Weevil", "Elephantiasis", "Fusarium", "Moko or Madurabiche", 
"Nematodes", "Red Spider Mite", "Scale Insect", "Scarab Beetle", 
"Whiteflies", "Yellow Sigatoka"))],
    crop == "Plantain" ~ colors_plantain_question5[match(answer_in_english, c("Banana Mosaic Disease", "Black Sigatoka", 
"Black Weevil", "Elephantiasis", "Fusarium", "Moko or Madurabiche", 
"Nematodes", "Red Spider Mite", "Scale Insect", "Scarab Beetle", 
"Whiteflies", "Yellow Sigatoka"))]
  ))
#Make the abbreviations
question_5_frequencies_by_depto <- question_5_frequencies_by_depto %>%
  mutate(cat_abbreviations = case_when(
    answer_in_english == "Banana Mosaic Disease" ~ "BMD",
    answer_in_english == "Black Sigatoka" ~ "BS",
    answer_in_english == "Black Weevil" ~ "BW",
    answer_in_english == "Elephantiasis" ~ "ELE",
    answer_in_english == "Fusarium" ~ "FUS",
    answer_in_english == "Moko or Madurabiche" ~ "MOKO",
    answer_in_english == "Nematodes" ~ "NEM",
    answer_in_english == "Red Spider Mite" ~ "RSM",
    answer_in_english == "Scale Insect" ~ "SI",
    answer_in_english == "Scarab Beetle" ~ "SB",
    answer_in_english == "Whiteflies" ~ "WF",
    answer_in_english == "Yellow Sigatoka" ~ "YS",
    TRUE ~ NA_character_  # Default case if none of the above conditions are met
  ))
#tree map
question_5_frequencies_by_depto<- select(question_5_frequencies_by_depto,!answer_in_english)
question_5_frequencies_by_depto<- question_5_frequencies_by_depto %>%
  group_by(crop,expert_in) %>%
  mutate(total= sum(n),
         mean= round((n/total)*100,0))
question_5_frequencies_by_depto<- select(question_5_frequencies_by_depto,!n & !total)
#put it in long format
question_5_frequencies_by_depto <- question_5_frequencies_by_depto %>% 
  uncount(weights = mean, .id = "id")
#only select top 5 for the tree map
  #for banana
  question_5_frequencies_by_depto_banana<- filter(question_5_frequencies_by_depto, cat_abbreviations %in% question_5_frequencies$codes[which(question_5_frequencies$h2=="Banana")], crop=="Banana")
  #for plantain
  question_5_frequencies_by_depto_plantain<- filter(question_5_frequencies_by_depto, cat_abbreviations %in% question_5_frequencies$codes[which(question_5_frequencies$h2=="Plantain")], crop=="Plantain")
#merge both data set
question_5_frequencies_by_depto<- rbind(question_5_frequencies_by_depto_banana, question_5_frequencies_by_depto_plantain)
#tree map only plantain
tree_map_3(data = filter(question_5_frequencies_by_depto, crop=="Plantain"), "crop", "expert_in", "cat_abbreviations", "group_color", title="Top 5 Most Limiting Pests and Diseases in Plantains by Department")
#tree map only banana
tree_map_3(data = filter(question_5_frequencies_by_depto, crop=="Banana"), "crop", "expert_in", "cat_abbreviations", "group_color", title="Top 5 Most Limiting Pests and Diseases in Bananas by Department")
#Additional top 5 diseases and pest in Colombia
q5_top_5Colobia<- count_elements_by_group(question_5, "answer_in_english", c("answer_in_english"))
q5_top_5Colobia$mean<-round((q5_top_5Colobia$n/sum(q5_top_5Colobia$n)*100),0)
q5_top_5Colobia$background_color<- "white"
q5_top_5Colobia$color<- palette_soil(length(unique(q5_top_5Colobia$answer_in_english)))
#Make abbreviations 
q5_top_5Colobia <- q5_top_5Colobia %>%
  mutate(cat_abbreviations = case_when(
    answer_in_english == "Banana Mosaic Disease" ~ "BMD",
    answer_in_english == "Black Sigatoka" ~ "BS",
    answer_in_english == "Black Weevil" ~ "BW",
    answer_in_english == "Elephantiasis" ~ "ELE",
    answer_in_english == "Fusarium" ~ "FUS",
    answer_in_english == "Moko or Madurabiche" ~ "MOKO",
    answer_in_english == "Nematodes" ~ "NEM",
    answer_in_english == "Red Spider Mite" ~ "RSM",
    answer_in_english == "Scale Insect" ~ "SI",
    answer_in_english == "Scarab Beetle" ~ "SB",
    answer_in_english == "Whiteflies" ~ "WF",
    answer_in_english == "Yellow Sigatoka" ~ "YS",
    TRUE ~ NA_character_  # Default case if none of the above conditions are met
  ))
#sort the results by mean
q5_top_5Colobia <- q5_top_5Colobia %>%
  arrange(desc(mean))
#Finding the order of the levels
q5_top_5Colobia$cat_abbreviations<- factor(q5_top_5Colobia$cat_abbreviations, levels = q5_top_5Colobia$cat_abbreviations)
#plot most important pest and diseases
bar_plot_banana_plantain(data_set = q5_top_5Colobia,x_lab = "Pest and Diseases", y_lab ="Proportion of Experts (%)", title = "
Important Pests and Diseases in Musaceae Identified by Experts" ,x = "cat_abbreviations", y = "mean", alpha = 0.3,background_color = "background_color", bar_color = "color",proportional_limit =1.10)

```

## Question 6: `r unique(individual_surveys$question_in_english)[6]`

```{r question_6}
#selecting only question 6----
  question_6<- filter(individual_surveys, question_number==6)
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_6 <- question_6 %>%
  mutate(answer_in_english = case_when(
    answer_in_english == "Very Low Knowledge,Low Knowledge,Moderate Knowledge" ~ "Low Knowledge",
    answer_in_english == "Very Low Knowledge,Low Knowledge" ~ "Low Knowledge",
    answer_in_english == "Moderate Knowledge,High Knowledge" ~ "Moderate Knowledge",
    TRUE ~ answer_in_english  # This keeps all other values as they are
  ))
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_6 <- question_6 %>%
  mutate(numeric_answer = case_when(
    numeric_answer == "2,3,4" ~ "3",
    numeric_answer == "2,3" ~ "2.5",
    numeric_answer == "4,5" ~ "4.5",
    TRUE ~ numeric_answer  # This keeps all other values as they are
  ))
  #organizing the order of the levels
  question_6$answer_in_english<- factor(question_6$answer_in_english,levels = 
c("No Knowledge", "Very Low Knowledge", "Low Knowledge", "Moderate Knowledge", "High Knowledge", "Very High Knowledge"))
 #average by department (weighted and no weighted)
    #not weighted 
    avg_dept_no_weight<- question_6 %>%
      group_by(crop, expert_in) %>%
      mutate(numeric_answer=as.numeric(numeric_answer))%>%
      mutate(average= mean(numeric_answer)) %>%
      group_by(average)%>%
      mutate(ave_cat= category_by_mean_by_question(6, mean =average, language = "en"))
    #weighted 
    add_weights_question6<- inner_join(question_6, experience_of_each_expert)
    avg_dept_weight<- add_weights_question6 %>%
      group_by(crop, expert_in) %>%
      mutate(numeric_answer=as.numeric(numeric_answer))%>%
      mutate(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
      group_by(average_weighted)%>%
      mutate(ave_cat_weighted= category_by_mean_by_question(6, mean =average_weighted, language = "en"))

    #finding the levels per crop (banana and plantain) no weighted
  levels_per_crop_Q6_no_weighted<- tapply(avg_dept_no_weight$ave_cat, avg_dept_no_weight$crop, function(x){(unique(x))})
  #generating a ramp palette according to the number of levels per crop
  colors_banana_question6_no_weighted<- palette_banana(length(levels_per_crop_Q6_no_weighted$Banana))
  colors_plantain_question6_no_weighted<- palette_plantain(length(levels_per_crop_Q6_no_weighted$Plantain))
  #Assigning the colors by crop and answer
    # Define the knowledge levels for each crop
    levels_by_crop_no_weighted <- list(
      Banana = levels_per_crop_Q6_no_weighted$Banana,
      Plantain = levels_per_crop_Q6_no_weighted$Plantain
    )
    # Define the colors for each crop
    colors_by_crop_no_weighted <- list(
      Banana = colors_banana_question6_no_weighted, # Replace with actual color vector for Banana
      Plantain = colors_plantain_question6_no_weighted # Replace with actual color vector for Plantain
      # Add more crops and their color vectors here if needed
    )
    # Now call the function
    avg_dept_no_weight <- assign_group_color(data = avg_dept_no_weight, levels_by_crop = levels_by_crop_no_weighted, colors_by_crop = colors_by_crop_no_weighted,answer_col =  "ave_cat")
    
    #finding the levels per crop (banana and plantain) weighted
  levels_per_crop_Q6_weighted<- tapply(avg_dept_weight$ave_cat_weighted, avg_dept_weight$crop, function(x){(unique(x))})
  #generating a ramp palette according to the number of levels per crop
  colors_banana_question6_weighted<- palette_banana(length(levels_per_crop_Q6_weighted$Banana))
  colors_plantain_question6_weighted<- palette_plantain(length(levels_per_crop_Q6_weighted$Plantain))
  #Assigning the colors by crop and answer
    # Define the knowledge levels for each crop
    levels_by_crop_weighted <- list(
      Banana = levels_per_crop_Q6_weighted$Banana,
      Plantain = levels_per_crop_Q6_weighted$Plantain
    )
    # Define the colors for each crop
    colors_by_crop_weighted <- list(
      Banana = colors_banana_question6_weighted, # Replace with actual color vector for Banana
      Plantain = colors_plantain_question6_weighted # Replace with actual color vector for Plantain
      # Add more crops and their color vectors here if needed
    )
    # Now call the function
    avg_dept_weight <- assign_group_color(data = avg_dept_weight, levels_by_crop = levels_by_crop_weighted, colors_by_crop = colors_by_crop_weighted,answer_col =  "ave_cat_weighted")

#tree map
tree_map_3(data = avg_dept_no_weight, "crop", "expert_in", "ave_cat", "group_color", title="No weighted")
tree_map_3(data = avg_dept_weight, "crop", "expert_in", "ave_cat_weighted", "group_color", title="weighted")
#Map of Colombia  
  #non weighted
  nal_q6_no_weight<- question_6 %>%
  group_by(expert_in) %>%
  summarise(average= mean(as.numeric(numeric_answer))) %>% 
  group_by(average) %>%
  mutate(cat= category_by_mean_by_question(language = "en", mean = average, question =6)) %>%
  rename(EKE.expert.in=expert_in)
  nal_q6_no_weight<- full_join(codes_department, nal_q6_no_weight)
  nal_q6_no_weight$cat<- factor(nal_q6_no_weight$cat, levels = c("Very Low Knowledge","Low Knowledge", "Moderate Knowledge", "High Knowledge", "Very High Knowledge", NA))
  colmap(departamentos, data = nal_q6_no_weight, data_id = "id_depto", var = "cat")+
    scale_fill_manual(values = palette_soil(length(unique(nal_q6_no_weight$cat))-1), na.value = "#eeeeee")
# weighted
  nal_q6_weight<- inner_join(question_6, experience_of_each_expert)
  nal_q6_weight<- nal_q6_weight %>%
  group_by(expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer)) %>%
  summarise(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted) %>%
  mutate(cat_weighted= category_by_mean_by_question(language = "en", mean = average_weighted, question =6)) %>%
  rename(EKE.expert.in=expert_in)
  
  nal_q6_weight<- full_join(codes_department, nal_q6_weight)
  nal_q6_weight$cat_weighted<- factor(nal_q6_weight$cat_weighted, levels = c("Very Low Knowledge","Low Knowledge", "Moderate Knowledge", "High Knowledge", "Very High Knowledge", NA))
  colmap(departamentos, data = nal_q6_weight, data_id = "id_depto", var = "cat_weighted")+
    scale_fill_manual(values = palette_soil(length(unique(nal_q6_weight$cat))-1), na.value = "#eeeeee")


```

## Question 7: `r unique(individual_surveys$question_in_english)[7]`

```{r question_7}
#selecting only question 7----
question_7<- filter(individual_surveys, question_number==7)
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_7 <- question_7 %>%
  mutate(answer_in_english = case_when(
    answer_in_english == "Somewhat Likely,Likely" ~ "Likely",
    answer_in_english == "Extremely Unlikely,Unlikely" ~ "Very Unlikely",
    TRUE ~ answer_in_english  # This keeps all other values as they are
  ))
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_7 <- question_7 %>%
  mutate(numeric_answer = case_when(
    numeric_answer == "1,3" ~ "2",
    numeric_answer == "6,7" ~ "6.5",
    TRUE ~ numeric_answer  # This keeps all other values as they are
  ))
#organizing the order of the levels
question_7$answer_in_english<- factor(question_7$answer_in_english,levels = c("Extremely Unlikely", "Very Unlikely", "Somewhat Unlikely", 
                                                                              "Unlikely", "Neutral", "Somewhat Likely", "Likely", 
                                                                              "Very Likely", "Extremely Likely"))
#average by department (weighted and no weighted)
#not weighted 
avg_dept_no_weight_7<- question_7 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average= mean(numeric_answer)) %>%
  group_by(average)%>%
  mutate(ave_cat= category_by_mean_by_question(7, mean =average, language = "en"))
#weighted 
add_weights_question_7<- inner_join(question_7, experience_of_each_expert)
avg_dept_weight_7<- add_weights_question_7 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted)%>%
  mutate(ave_cat_weighted= category_by_mean_by_question(7, mean =average_weighted, language = "en"))

#finding the levels per crop (banana and plantain) no weighted
levels_per_crop_Q7_no_weighted<- tapply(avg_dept_no_weight_7$ave_cat, avg_dept_no_weight_7$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question7_no_weighted<- palette_banana(length(levels_per_crop_Q7_no_weighted$Banana))
colors_plantain_question7_no_weighted<- palette_plantain(length(levels_per_crop_Q7_no_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_no_weighted_7 <- list(
  Banana = levels_per_crop_Q7_no_weighted$Banana,
  Plantain = levels_per_crop_Q7_no_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_no_weighted_7 <- list(
  Banana = colors_banana_question7_no_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question7_no_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_no_weight_7 <- assign_group_color(data = avg_dept_no_weight_7, levels_by_crop = levels_by_crop_no_weighted_7, colors_by_crop = colors_by_crop_no_weighted_7,answer_col =  "ave_cat")

#finding the levels per crop (banana and plantain) weighted
levels_per_crop_Q7_weighted<- tapply(avg_dept_weight_7$ave_cat_weighted, avg_dept_weight_7$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question7_weighted<- palette_banana(length(levels_per_crop_Q7_weighted$Banana))
colors_plantain_question7_weighted<- palette_plantain(length(levels_per_crop_Q7_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_weighted_7 <- list(
  Banana = levels_per_crop_Q7_weighted$Banana,
  Plantain = levels_per_crop_Q7_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_weighted_7 <- list(
  Banana = colors_banana_question7_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question7_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_weight_7 <- assign_group_color(data = avg_dept_weight_7, levels_by_crop = levels_by_crop_weighted_7, colors_by_crop = colors_by_crop_weighted_7,answer_col =  "ave_cat_weighted")

#tree map
tree_map_3(data = avg_dept_no_weight_7, "crop", "expert_in", "ave_cat", "group_color", title="No weighted")
tree_map_3(data = avg_dept_weight_7, "crop", "expert_in", "ave_cat_weighted", "group_color", title="weighted")
#Map of Colombia  
#non weighted
nal_q7_no_weight<- question_7 %>% group_by(expert_in) %>%
  summarise(average= mean(as.numeric(numeric_answer))) %>%
  group_by(average) %>% 
  mutate(cat= category_by_mean_by_question(language = "en", mean = average, question =7))%>%
  rename(EKE.expert.in = expert_in)
  
nal_q7_no_weight<- full_join(codes_department, nal_q7_no_weight)




nal_q7_no_weight$cat<- factor(nal_q7_no_weight$cat, levels = c("Extremely Unlikely", "Very Unlikely", "Somewhat Unlikely", 
                                                                        "Unlikely", "Neutral", "Somewhat Likely", "Likely", 
                                                                        "Very Likely", "Extremely Likely"))
colmap(departamentos, data = nal_q7_no_weight, data_id = "id_depto", var = "cat")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q7_no_weight$cat))-1), na.value = "#eeeeee")
# weighted
nal_q7_weight<- inner_join(question_7, experience_of_each_expert)
nal_q7_weight<- nal_q7_weight %>%
  group_by(expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer)) %>%
  summarise(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted) %>%
  mutate(cat_weighted= category_by_mean_by_question(language = "en", mean = average_weighted, question =7)) %>%
  rename(EKE.expert.in=expert_in)

nal_q7_weight<- full_join(codes_department, nal_q7_weight)
nal_q7_weight$cat_weighted<- factor(nal_q7_weight$cat_weighted, levels = c("Extremely Unlikely", "Very Unlikely", "Somewhat Unlikely", 
                                                                           "Unlikely", "Neutral", "Somewhat Likely", "Likely", 
                                                                           "Very Likely", "Extremely Likely"))
colmap(departamentos, data = nal_q7_weight, data_id = "id_depto", var = "cat_weighted")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q7_weight$cat))-1), na.value = "#eeeeee")

```

## Question 8: `r unique(individual_surveys$question_in_english)[8]`

```{r question_8}
#selecting only question 8----
question_8<- filter(individual_surveys, question_number==8)
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_8 <- question_8 %>%
  mutate(answer_in_english = case_when(
    answer_in_english == "Slightly Effective,Moderately Effective" ~ "Moderately Effective",
    TRUE ~ answer_in_english  # This keeps all other values as they are
  ))
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_8 <- question_8 %>%
  mutate(numeric_answer = case_when(
    numeric_answer == "2,3" ~ "2.5",
    TRUE ~ numeric_answer  # This keeps all other values as they are
  ))
#organizing the order of the levels
question_8$answer_in_english<- factor(question_8$answer_in_english,levels = c("Not Effective", "Slightly Effective", "Moderately Effective", "Very Effective", "Extremely Effective"))
#average by department (weighted and no weighted)
#not weighted 
avg_dept_no_weight_8<- question_8 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average= mean(numeric_answer)) %>%
  group_by(average)%>%
  mutate(ave_cat= category_by_mean_by_question(8, mean =average, language = "en"))
#weighted 
add_weights_question_8<- inner_join(question_8, experience_of_each_expert)
avg_dept_weight_8<- add_weights_question_8 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted)%>%
  mutate(ave_cat_weighted= category_by_mean_by_question(8, mean =average_weighted, language = "en"))

#finding the levels per crop (banana and plantain) no weighted
levels_per_crop_Q8_no_weighted<- tapply(avg_dept_no_weight_8$ave_cat, avg_dept_no_weight_8$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question8_no_weighted<- palette_banana(length(levels_per_crop_Q8_no_weighted$Banana))
colors_plantain_question8_no_weighted<- palette_plantain(length(levels_per_crop_Q8_no_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_no_weighted_8 <- list(
  Banana = levels_per_crop_Q8_no_weighted$Banana,
  Plantain = levels_per_crop_Q8_no_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_no_weighted_8 <- list(
  Banana = colors_banana_question8_no_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question8_no_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_no_weight_8 <- assign_group_color(data = avg_dept_no_weight_8, levels_by_crop = levels_by_crop_no_weighted_8, colors_by_crop = colors_by_crop_no_weighted_8,answer_col =  "ave_cat")

#finding the levels per crop (banana and plantain) weighted
levels_per_crop_Q8_weighted<- tapply(avg_dept_weight_8$ave_cat_weighted, avg_dept_weight_8$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question8_weighted<- palette_banana(length(levels_per_crop_Q8_weighted$Banana))
colors_plantain_question8_weighted<- palette_plantain(length(levels_per_crop_Q8_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_weighted_8 <- list(
  Banana = levels_per_crop_Q8_weighted$Banana,
  Plantain = levels_per_crop_Q8_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_weighted_8 <- list(
  Banana = colors_banana_question8_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question8_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_weight_8 <- assign_group_color(data = avg_dept_weight_8, levels_by_crop = levels_by_crop_weighted_8, colors_by_crop = colors_by_crop_weighted_8,answer_col =  "ave_cat_weighted")

#tree map
tree_map_3(data = avg_dept_no_weight_8, "crop", "expert_in", "ave_cat", "group_color", title="No weighted")
tree_map_3(data = avg_dept_weight_8, "crop", "expert_in", "ave_cat_weighted", "group_color", title="weighted")
#Map of Colombia  
#non weighted
nal_q8_no_weight<- question_8 %>% group_by(expert_in) %>%
  summarise(average= mean(as.numeric(numeric_answer))) %>%
  group_by(average) %>% 
  mutate(cat= category_by_mean_by_question(language = "en", mean = average, question =8))%>%
  rename(EKE.expert.in = expert_in)
  
nal_q8_no_weight<- full_join(codes_department, nal_q8_no_weight)
nal_q8_no_weight$cat<- factor(nal_q8_no_weight$cat, levels = c("Not Effective", "Slightly Effective", "Moderately Effective", "Very Effective", "Extremely Effective"))
colmap(departamentos, data = nal_q8_no_weight, data_id = "id_depto", var = "cat")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q8_no_weight$cat))-1), na.value = "#eeeeee")
# weighted
nal_q8_weight<- inner_join(question_8, experience_of_each_expert)
nal_q8_weight<- nal_q8_weight %>%
  group_by(expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer)) %>%
  summarise(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted) %>%
  mutate(cat_weighted= category_by_mean_by_question(language = "en", mean = average_weighted, question =8)) %>%
  rename(EKE.expert.in=expert_in)

nal_q8_weight<- full_join(codes_department, nal_q8_weight)
nal_q8_weight$cat_weighted<- factor(nal_q8_weight$cat_weighted, levels = c("Not Effective", "Slightly Effective", "Moderately Effective", "Very Effective", "Extremely Effective"))
colmap(departamentos, data = nal_q8_weight, data_id = "id_depto", var = "cat_weighted")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q8_weight$cat))-1), na.value = "#eeeeee")
```

## Question 9: `r unique(individual_surveys$question_in_english)[9]`

```{r question_9}
#selecting only question 9----
question_9<- filter(individual_surveys, question_number==9)
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_9 <- question_9 %>%
  mutate(answer_in_english = case_when(
    answer_in_english == "Adequate,Good" ~ "Good",
    TRUE ~ answer_in_english  # This keeps all other values as they are
  ))
#some experts selected more than one option in this question so an intermediate answer is chosen 
question_9 <- question_9 %>%
  mutate(numeric_answer = case_when(
    numeric_answer == "4,5" ~ "4.5",
    TRUE ~ numeric_answer  # This keeps all other values as they are
  ))
#organizing the order of the levels
question_9$answer_in_english<- factor(question_9$answer_in_english,levels = c("No Training or Information Available", "Poor", "Fair", "Adequate", "Good", "Excellent", "Unsure"))
question_9<- filter(question_9, answer_in_english!="Unsure")
#average by department (weighted and no weighted)
#not weighted 
avg_dept_no_weight_9<- question_9 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average= mean(numeric_answer)) %>%
  group_by(average)%>%
  mutate(ave_cat= category_by_mean_by_question(9, mean =average, language = "en"))
#weighted 
add_weights_question_9<- inner_join(question_9, experience_of_each_expert)
avg_dept_weight_9<- add_weights_question_9 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted)%>%
  mutate(ave_cat_weighted= category_by_mean_by_question(9, mean =average_weighted, language = "en"))

#finding the levels per crop (banana and plantain) no weighted
levels_per_crop_Q9_no_weighted<- tapply(avg_dept_no_weight_9$ave_cat, avg_dept_no_weight_9$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question9_no_weighted<- palette_banana(length(levels_per_crop_Q9_no_weighted$Banana))
colors_plantain_question9_no_weighted<- palette_plantain(length(levels_per_crop_Q9_no_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_no_weighted_9 <- list(
  Banana = levels_per_crop_Q9_no_weighted$Banana,
  Plantain = levels_per_crop_Q9_no_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_no_weighted_9 <- list(
  Banana = colors_banana_question9_no_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question9_no_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_no_weight_9 <- assign_group_color(data = avg_dept_no_weight_9, levels_by_crop = levels_by_crop_no_weighted_9, colors_by_crop = colors_by_crop_no_weighted_9,answer_col =  "ave_cat")

#finding the levels per crop (banana and plantain) weighted
levels_per_crop_Q9_weighted<- tapply(avg_dept_weight_9$ave_cat_weighted, avg_dept_weight_9$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question9_weighted<- palette_banana(length(levels_per_crop_Q9_weighted$Banana))
colors_plantain_question9_weighted<- palette_plantain(length(levels_per_crop_Q9_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_weighted_9 <- list(
  Banana = levels_per_crop_Q9_weighted$Banana,
  Plantain = levels_per_crop_Q9_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_weighted_9 <- list(
  Banana = colors_banana_question9_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question9_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_weight_9 <- assign_group_color(data = avg_dept_weight_9, levels_by_crop = levels_by_crop_weighted_9, colors_by_crop = colors_by_crop_weighted_9,answer_col =  "ave_cat_weighted")

#tree map
tree_map_3(data = avg_dept_no_weight_9, "crop", "expert_in", "ave_cat", "group_color", title="No weighted")
tree_map_3(data = avg_dept_weight_9, "crop", "expert_in", "ave_cat_weighted", "group_color", title="weighted")
#Map of Colombia  
#non weighted
nal_q9_no_weight<- question_9 %>% group_by(expert_in) %>%
  summarise(average= mean(as.numeric(numeric_answer))) %>%
  group_by(average) %>% 
  mutate(cat= category_by_mean_by_question(language = "en", mean = average, question =9))%>%
  rename(EKE.expert.in = expert_in)
  
nal_q9_no_weight<- full_join(codes_department, nal_q9_no_weight)
nal_q9_no_weight$cat<- factor(nal_q9_no_weight$cat, levels = c("No Training or Information Available", "Poor", "Fair", "Adequate", "Good", "Excellent", "Unsure"))
colmap(departamentos, data = nal_q9_no_weight, data_id = "id_depto", var = "cat")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q9_no_weight$cat))-1), na.value = "#eeeeee")
# weighted
nal_q9_weight<- inner_join(question_9, experience_of_each_expert)
nal_q9_weight<- nal_q9_weight %>%
  group_by(expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer)) %>%
  summarise(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted) %>%
  mutate(cat_weighted= category_by_mean_by_question(language = "en", mean = average_weighted, question =9)) %>%
  rename(EKE.expert.in=expert_in)

nal_q9_weight<- full_join(codes_department, nal_q9_weight)
nal_q9_weight$cat_weighted<- factor(nal_q9_weight$cat_weighted, levels = c("No Training or Information Available", "Poor", "Fair", "Adequate", "Good", "Excellent", "Unsure"))
colmap(departamentos, data = nal_q9_weight, data_id = "id_depto", var = "cat_weighted")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q9_weight$cat))-1), na.value = "#eeeeee")
```

## Question 13: `r unique(individual_surveys$question_in_english)[13]`

```{r questoin_13}

#selecting only question 13----
question_13<- filter(individual_surveys, question_number==13)
#organizing the order of the levels
question_13$answer_in_english<- factor(question_13$answer_in_english,levels = c("Every 1 or 2 cycles", "Every 3 to 4 cycles", "Every 5 to 6 cycles",  "Every 7 or more cycles", "Not Sure"))
question_13<- filter(question_13, answer_in_english!="Not Sure")
#average by department (weighted and no weighted)
#not weighted 
avg_dept_no_weight_13<- question_13 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average= mean(numeric_answer)) %>%
  group_by(average)%>%
  mutate(ave_cat= category_by_mean_by_question(13, mean =average, language = "en"))
#weighted 
add_weights_question_13<- inner_join(question_13, experience_of_each_expert)
avg_dept_weight_13<- add_weights_question_13 %>%
  group_by(crop, expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer))%>%
  mutate(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted)%>%
  mutate(ave_cat_weighted= category_by_mean_by_question(13, mean =average_weighted, language = "en"))

#finding the levels per crop (banana and plantain) no weighted
levels_per_crop_Q13_no_weighted<- tapply(avg_dept_no_weight_13$ave_cat, avg_dept_no_weight_13$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question13_no_weighted<- palette_banana(length(levels_per_crop_Q13_no_weighted$Banana))
colors_plantain_question13_no_weighted<- palette_plantain(length(levels_per_crop_Q13_no_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_no_weighted_13 <- list(
  Banana = levels_per_crop_Q13_no_weighted$Banana,
  Plantain = levels_per_crop_Q13_no_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_no_weighted_13 <- list(
  Banana = colors_banana_question13_no_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question13_no_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_no_weight_13 <- assign_group_color(data = avg_dept_no_weight_13, levels_by_crop = levels_by_crop_no_weighted_13, colors_by_crop = colors_by_crop_no_weighted_13,answer_col =  "ave_cat")

#finding the levels per crop (banana and plantain) weighted
levels_per_crop_Q13_weighted<- tapply(avg_dept_weight_13$ave_cat_weighted, avg_dept_weight_13$crop, function(x){(unique(x))})
#generating a ramp palette according to the number of levels per crop
colors_banana_question13_weighted<- palette_banana(length(levels_per_crop_Q13_weighted$Banana))
colors_plantain_question13_weighted<- palette_plantain(length(levels_per_crop_Q13_weighted$Plantain))
#Assigning the colors by crop and answer
# Define the knowledge levels for each crop
levels_by_crop_weighted_13 <- list(
  Banana = levels_per_crop_Q13_weighted$Banana,
  Plantain = levels_per_crop_Q13_weighted$Plantain
)
# Define the colors for each crop
colors_by_crop_weighted_13 <- list(
  Banana = colors_banana_question13_weighted, # Replace with actual color vector for Banana
  Plantain = colors_plantain_question13_weighted # Replace with actual color vector for Plantain
  # Add more crops and their color vectors here if needed
)
# Now call the function
avg_dept_weight_13 <- assign_group_color(data = avg_dept_weight_13, levels_by_crop = levels_by_crop_weighted_13, colors_by_crop = colors_by_crop_weighted_13,answer_col =  "ave_cat_weighted")

#tree map
tree_map_3(data = avg_dept_no_weight_13, "crop", "expert_in", "ave_cat", "group_color", title="No weighted")
tree_map_3(data = avg_dept_weight_13, "crop", "expert_in", "ave_cat_weighted", "group_color", title="weighted")
#Map of Colombia  
#non weighted
nal_q13_no_weight<- question_13 %>% group_by(expert_in) %>%
  summarise(average= mean(as.numeric(numeric_answer))) %>%
  group_by(average) %>% 
  mutate(cat= category_by_mean_by_question(language = "en", mean = average, question =13))%>%
  rename(EKE.expert.in = expert_in)
  
nal_q13_no_weight<- full_join(codes_department, nal_q13_no_weight)
nal_q13_no_weight$cat<- factor(nal_q13_no_weight$cat, levels = c("Every 1 or 2 cycles", "Every 3 to 4 cycles", "Every 5 to 6 cycles",  "Every 7 or more cycles", "Not Sure"))
colmap(departamentos, data = nal_q13_no_weight, data_id = "id_depto", var = "cat")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q13_no_weight$cat))-1), na.value = "#eeeeee")
# weighted
nal_q13_weight<- inner_join(question_13, experience_of_each_expert)
nal_q13_weight<- nal_q13_weight %>%
  group_by(expert_in) %>%
  mutate(numeric_answer=as.numeric(numeric_answer)) %>%
  summarise(average_weighted= weighted.mean(numeric_answer, weights)) %>% 
  group_by(average_weighted) %>%
  mutate(cat_weighted= category_by_mean_by_question(language = "en", mean = average_weighted, question =13)) %>%
  rename(EKE.expert.in=expert_in)

nal_q13_weight<- full_join(codes_department, nal_q13_weight)
nal_q13_weight$cat_weighted<- factor(nal_q13_weight$cat_weighted, levels = c("Every 1 or 2 cycles", "Every 3 to 4 cycles", "Every 5 to 6 cycles",  "Every 7 or more cycles", "Not Sure"))
colmap(departamentos, data = nal_q13_weight, data_id = "id_depto", var = "cat_weighted")+
  scale_fill_manual(values = palette_soil(length(unique(nal_q13_weight$cat))-1), na.value = "#eeeeee")
```