---
title: "R Notebook"
output: html_notebook
---

## Network analysis:

This section perform the analysis of the banana and plantain planting meterial and stakeholder movement network.

### Preparing the data set for the network analysis

#### Network of planting material

```{r}
# Upload the network of panting material from github
net_plant_mat<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/DATA_MAP_Network_of_Planting_Material_2023_12_08.csv", header = TRUE, row.names = 1)
head(net_plant_mat)
```

The data set has more than one category by cell. First we will check what are all the categories that are in the data set. They shoul be: BC, BN, PC, PN for banana registerd and non-registered and plantain registered and non-registered, respectively.

```{r}
# Unlist the data to get all entries in a single vector
all_values <- unlist(net_plant_mat)
# Filter out the zeroes if they are not needed
all_values <- all_values[all_values != "0"]
# Split the comma-separated values and unlist again
all_values_split <- unlist(strsplit(all_values, ","))
# Get unique values
unique_values <- unique(all_values_split)
# Display the unique values
print(unique_values)
```

The data set is correctly formatted. Now we will need to create a binary matrix for each category (four in total --\> "BC" "BN" "PN" "PC"). We will create a matrix of the same size as the original data set, but with 1s and 0s. 1s will indicate that the category is present in the cell, and 0s will indicate that it is not.

```{r}
# Get the dimensions of the original data
n_rows <- nrow(net_plant_mat)
n_cols <- ncol(net_plant_mat)
# Initialize a list to store the binary matrices for each category
binary_matrices <- list()
# Define the four categories
categories <- c("BC", "BN", "PN", "PC")
# Loop through each category
for (category in categories) {
  # Initialize an empty matrix of the same size as net_plant_mat with 0s
  binary_matrix <- matrix(0, nrow = n_rows, ncol = n_cols)
  # Loop through each cell in the data frame
  for (i in 1:n_rows) {
    for (j in 1:n_cols) {
      # Check if the category is present in the current cell
      if (grepl(category, net_plant_mat[i, j])) {
        binary_matrix[i, j] <- 1
      }
      
    }
  }
  # Store the binary matrix in the list with the category name as the key
  binary_matrices[[category]] <- binary_matrix
  # Convert the matrix to a data frame and assign row and column names
  binary_df <- as.data.frame(binary_matrix)
  rownames(binary_df) <- rownames(net_plant_mat)
  colnames(binary_df) <- colnames(net_plant_mat)
  # Create the file name using the specified format
  file_name <- paste0("C:/Users/jrobl/OneDrive - University of Florida/Documents/TA_2023_R2M/R2M_Colombia_banana_and_plantain/Data/DATA_MAP_Net_", category, ".csv")
  # Write the data frame to a CSV file
  write.csv(binary_df, file_name, row.names = TRUE)
}
```

Check that the unique categories between the binary data sets and the original data set is the same.

```{r}
# Unlist the data to get all entries in a single vector
all_values <- unlist(net_plant_mat)
# Filter out the zeroes if they are not needed
all_values <- all_values[all_values != "0"]
# Split the comma-separated values and unlist again
all_values_split <- unlist(strsplit(all_values, ","))
# Get unique values
unique_original_categories <- unique(all_values_split)
# Loop through each category and check if the binary matrix is correct
for (category in categories) {
  # Get the corresponding binary matrix
  binary_matrix <- binary_matrices[[category]]
  # Extract the positions where there is a 1 in the binary matrix
  positions <- which(binary_matrix == 1, arr.ind = TRUE)
  # Check if all positions in the original matrix contain the correct category
  for (pos in 1:nrow(positions)) {
    row <- positions[pos, 1]
    col <- positions[pos, 2]
    # Verify the category is in the original matrix at these positions
    if (!grepl(category, net_plant_mat[row, col])) {
      stop(paste("Mismatch found for category", category, "at position (", row, ",", col, ")"))
    }
  }
}
# If the script completes without stopping, it means all checks passed
cat("All binary matrices correctly reflect the categories in the original data set.\n")
```

The four data sets will be used later for the network analysis. Names of the files:

-   DATA_MAP_Net_BC.csv
-   DATA_MAP_Net_BN.csv
-   DATA_MAP_Net_PN.csv
-   DATA_MAP_Net_PC.csv

#### Data set of the planted area by department in the last three years

This data set was taken from AGRONET. It is a data set that contains the planted area of banana and plantain by department in the last three years (from 2020 to 2022, needs to be filtered). The data set for banana is the sum of the information provided in this web page for these crops: BANANO y BANANO EXPORTACION . The data set for plantain is the sum of the information provided in this web page for these crops: PLATANO, and PLATANO EXPORTACION. This data set will be used to calculate the absolute link weihgts of the plantain material network. For that we assume that the *demand of planting material* of each department is proportional to cum sum of the change of planted area in the last three years. We have assumption that if the cum sum of the change of planted area is negative there is a superhabit of planting material in that department and therefore the demand is very small *(set as 1 ha)*. If it is positive there is a deficit of planting material. The steps to calculate the cum sum of the change of planted area area:

**Step 1: Calculate Yearly Area Changes**

$$
\Delta A_{i,t} = A_{i,t+1} - A_{i,t} \quad \text{for } t = 2020 \text{ to } 2021
$$

**Step 2: Calculate Cumulative Area Change**

$$
\Delta A_{i} = \sum_{t=2020}^{2021} \Delta A_{i,t}
$$

```{r}
# Upload the data set from github
banana<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/SUP_AREA_Banano.csv", header = TRUE)
banana_exp<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/SUP_AREA_Banano_Exportacion.csv", header = TRUE)
plantain<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/SUP_AREA_Platano.csv", header = TRUE)
plantain_exp<- read.csv("https://raw.githubusercontent.com/jrobledob/R2M_Colombia_banana_and_plantain/main/Data/SUP_AREA_Platano_Exportacion.csv", header = TRUE)

#merge the banana data sets and the plantain data sets
banana<- rbind(banana, banana_exp)
plantain<- rbind(plantain, plantain_exp)
head(banana)
```

```{r}
head(plantain)
```

This data sets (banana and plantain) need to be process for the network analysis. The steps are:

-   Select only the cells which Año is either 2020, 2021 or 2022
-   Group by Departamento and sum the values of the planted area
-   Create a new column called "Cultivo" with the name of the crop (banana or plantain)
-   The years 2020, 2021 and 2022 will be independent columns
-   Calculate the delta years (2022-2021, 2021-2020) for the planted area in two new columns for each crop

```{r}
# Select only the cells which Año is either 2020, 2021 or 2022
banana<- banana[banana$Año %in% c(2020, 2021, 2022),]
plantain<- plantain[plantain$Año %in% c(2020, 2021, 2022),]
#Remove the ',' from the Area..ha. column and transform it to numeric
banana$Area..ha.<- as.numeric(gsub(",", "", banana$Area..ha.))
plantain$Area..ha.<- as.numeric(gsub(",", "", plantain$Area..ha.))
# Group by Departamento and Año and sum the values of the planted area (Area..ha.)
banana<- aggregate(banana$Area..ha., by = list(banana$Departamento, banana$Año), FUN = sum)
plantain<- aggregate(plantain$Area..ha., by = list(plantain$Departamento, plantain$Año), FUN = sum)
# Create a new column called "Cultivo" with the name of the crop (banana or plantain)
banana$Cultivo<- "Banana"
plantain$Cultivo<- "Plantain"
#rename the columns 
colnames(banana)<- c("Departamento", "Año", "Area..ha.", "Cultivo")
colnames(plantain)<- c("Departamento", "Año", "Area..ha.", "Cultivo")
#make the categories of the Año indepent columns
library(tidyr)
banana<- spread(banana, key = Año, value = Area..ha.)
plantain<- spread(plantain, key = Año, value = Area..ha.)
#Calculate the delta years (2022-2021, 2021-2020) for the planted area in two new columns for each crop
banana$delta_2022_2021<- banana$`2022`-banana$`2021`
banana$delta_2021_2020<- banana$`2021`-banana$`2020`
#caluculate the cum sum of the change of planted area
banana$delta_cumsum<- banana$delta_2022_2021+banana$delta_2021_2020
plantain$delta_2022_2021<- plantain$`2022`-plantain$`2021`
plantain$delta_2021_2020<- plantain$`2021`-plantain$`2020`
#caluculate the cum sum of the change of planted area
plantain$delta_cumsum<- plantain$delta_2022_2021+plantain$delta_2021_2020
# Make the cumsum of the negative values equal to 1
banana$delta_cumsum[banana$delta_cumsum<0]<- 1
plantain$delta_cumsum[plantain$delta_cumsum<0]<- 1
head(banana)
```

```{r}
head(plantain)
```

The `delta_cumsum` column is equivalent to the increase in planted area in the last three years. This area (in hectares) will be multiplied by the number of plants per hectare to get the demand for planting material in each department.

*The planting density for plantain ($\rho_{\text{plantain}}$) is 1600 plants per hectare, and for banana ($\rho_{\text{banana}}$) is 2700 plants per hectare.*

```{r}
# Define the planting densities
rho_plantain <- 1600
rho_banana <- 2700
# Calculate the demand for planting material in each department
banana$demanda<- banana$delta_cumsum*rho_banana
plantain$demanda<- plantain$delta_cumsum*rho_plantain
head(banana)
```
```{r}
head(plantain)
```

Next the data sets must be unified with the names of in the network of planting material data set (net_plant_mat). For that we will change the names in net_plant_mat to uppercase and remove the accents.


```{r}
# identify the names in the plantain data sets
dput(unique(plantain$Departamento))
```
```{r}
# identify the names in the banana data sets
dput(unique(banana$Departamento))
```


```{r}
# Identify the names in the network of planting material data set
dput(unique(colnames(net_plant_mat)))
```
```{r}
# Create a named vector where the names correspond to the format in net_plant_mat
name_mapping_plantain <- c(
  "AMAZONAS" = "Amazonas", 
  "ANTIOQUIA" = "Antioquia", 
  "ARAUCA" = "Arauca", 
  "ATLANTICO" = "Atlántico", 
  "BOLIVAR" = "Bolívar", 
  "BOYACA" = "Boyacá", 
  "CALDAS" = "Caldas", 
  "CAQUETA" = "Caquetá", 
  "CASANARE" = "Casanare", 
  "CAUCA" = "Cauca", 
  "CESAR" = "Cesar", 
  "CHOCO" = "Chocó", 
  "CORDOBA" = "Córdoba", 
  "CUNDINAMARCA" = "Cundinamarca", 
  "GUAINIA" = "Guainía", 
  "GUAVIARE" = "Guaviare", 
  "HUILA" = "Huila", 
  "LA GUAJIRA" = "La.Guajira", 
  "MAGDALENA" = "Magdalena", 
  "META" = "Meta", 
  "NARIÑO" = "Nariño", 
  "NORTE DE SANTANDER" = "Norte.de.Santander", 
  "PUTUMAYO" = "Putumayo", 
  "QUINDIO" = "Quindío", 
  "RISARALDA" = "Risaralda", 
  "SAN ANDRES Y PROVIDENCIA" = "San.Andrés.y.Providencia", 
  "SANTANDER" = "Santander", 
  "SUCRE" = "Sucre", 
  "TOLIMA" = "Tolima", 
  "VALLE DEL CAUCA" = "Valle.del.Cauca", 
  "VAUPES" = "Vaupés", 
  "VICHADA" = "Vichada"
)
# Replace the names in plantain$Departamento using the mapping
plantain$Departamento <- name_mapping_plantain[plantain$Departamento]

# Check if all names have been replaced correctly
dput(unique(plantain$Departamento))
```
```{r}
# For bananas
name_mapping_bn <- c(
  "ANTIOQUIA" = "Antioquia", 
  "ATLANTICO" = "Atlántico", 
  "BOYACA" = "Boyacá", 
  "CALDAS" = "Caldas", 
  "CAQUETA" = "Caquetá", 
  "CASANARE" = "Casanare", 
  "CAUCA" = "Cauca", 
  "CESAR" = "Cesar", 
  "CHOCO" = "Chocó", 
  "CUNDINAMARCA" = "Cundinamarca", 
  "HUILA" = "Huila", 
  "LA GUAJIRA" = "La.Guajira", 
  "MAGDALENA" = "Magdalena", 
  "META" = "Meta", 
  "NARIÑO" = "Nariño", 
  "NORTE DE SANTANDER" = "Norte.de.Santander", 
  "PUTUMAYO" = "Putumayo", 
  "QUINDIO" = "Quindío", 
  "RISARALDA" = "Risaralda", 
  "SAN ANDRES Y PROVIDENCIA" = "San.Andrés.y.Providencia", 
  "SANTANDER" = "Santander",
  "TOLIMA" = "Tolima", 
  "VALLE DEL CAUCA" = "Valle.del.Cauca", 
  "VAUPES" = "Vaupés")
# Replace the names in banana$Departamento using the mapping
banana$Departamento <- name_mapping_bn[banana$Departamento]
# Check if all names have been replaced correctly
dput(unique(banana$Departamento))
```
Save the data set only with the name of the department and the demand of planting material.

# Problem!!! some values have NAs in the banana data set. We need to deal with them.


```{r}
# Select only the columns Departamento and demanda
banana<- banana[,c("Departamento", "demanda", "Cultivo")]
plantain<- plantain[,c("Departamento", "demanda",  "Cultivo")]
# Write the data frame to a CSV file
write.csv(banana, "C:/Users/jrobl/OneDrive - University of Florida/Documents/TA_2023_R2M/R2M_Colombia_banana_and_plantain/Data/DATA_Seed_demand_Banana.csv", row.names = FALSE)
write.csv(plantain, "C:/Users/jrobl/OneDrive - University of Florida/Documents/TA_2023_R2M/R2M_Colombia_banana_and_plantain/Data/DATA_Seed_demand_Plantain.csv", row.names = FALSE)
```


